---
title: "Final Project Data Memo"
author: "Louis Yang, Will Cichowski, Matthew Lorenz, Olivia Lee"
date: '2022-04-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Timeline

<br>

At this point in the project, our group, the Variance Acknowledgers, have settled on our dataset, which we will go in more depth about later in this memo. At this moment, we have been able to successfully read in our data to R, which is in a csv format. In the coming couple of weeks, we will be performing an EDA on this dataset, getting to know it better, and beginning to think more about the direction of our final project. After that, once we have met to talk about direction, we will begin the coding-heavy portions of our project and perform deeper analysis throughout the rest of the quarter.

<br>

## Overview of Dataset

<br>

Below is a skim of the dataset in its raw form, no cleaning performed on it yet.

<br>

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(skimr)
library(readr)
tinder <- read_csv('data/unprocessed/Tinder Millennial Match Rate Dataset.csv')
skim_without_charts(tinder)
```

<br>

This data is fairly small, with only 453 observations. We found it on Kaggle in a csv format. There is no missingness to the data.

<br>

It consists of a questionnaire that was answered by millennial aged people from various universities about their usage of the popular dating app, Tinder. Some of these questions, which are the columns of each dataset, are medium of use, such as mobile or web browser (currently Segment Type column), Name of the university (currently Segment description), whether or not they use Tinder (Answer), a count of how many matches they have gotten on Tinder, the percentage of matches they get when they swipe right, and a categorical variable of whether or not they succeeded in finding a relationship on Tinder. There is also a unique ID column for each observation. Most likely, we will be intending to predict an outcome variable of whether or not they achieved a relationship from Tinder. 

<br>

## Potential Research Questions
<br>
<br>

There are a variety of different research questions we can entertain given this dataset. The first, and most interesting to our group is: does usage lead to success? We can also model all of the variables to see what leads to a successful relationship. We can also build different models to see if the medium of usage leads to more success or less success in a relationship. I believe most of our models would be using forms of binary classification outcomes in order to measure success. Therefore, our metrics could be accuracy, precision, recall, and f1-measurement. I think f1 is probably the best metric to use in our case because of the imbalance in the dataset (over 70% have not had successful relationships).


## Potential Difficulties
<br>
<br>
The most obvious difficulty given this data set is the lack of complexity. The data set itself only has ~500 observations, and the only contains 6 usable variables. This will limit us in terms of cross validation, because if we created 10 folds out of 80% of our training data, we only are using 40 observations to train. This could result in each fold generating drastically different values. Further, having 6 usable variables limits our feature engineering. There is simply not that many combinations of new variables we can make. The most worrisome issue is the confounding in the data. For example, if someone has never used variable, obviously they never would have gotten a successful relationship from it. Moving forward, I am unsure if this is a dataset we want to use. 




